\vspace{-1.5em} \bf A general and efficient representation of ancestral recombination graphs

### 2023-10-16 paper.tex
yan's upates

Fully decoupling the general concept of an ARG from the coalescent
with recombination stochastic process is an important step.
While this has proven to be a useful and
robust
model~\citep{wakeley2012gene,bhaskar2014distortion,nelson2020accounting},
many modern datasets have properties that grossly
violate its assumptions.
For instance, several human datasets now consist of hundreds of thousands of
% TODO cite GeL paper
genomes~\citep{bycroft2018genome,karczewski2020mutational,tanjo2021practical},
and so sample size is substantially \emph{larger} than the
usually assumed $N_e$ values.
Agricultural datasets are an even more extreme departure from coalescent
assumptions, with hundreds of thousands of samples embedded in
multi-generational pedigrees~\citep{hayes20191000,Ros-Freixedes2020}
and effective population sizes of 100 and even
less~\citep{MacLeod2013,Makanjuola2020,Hall2016,Porcnic2016}.
Even if sufficiently complex demographic models~\citep{gower2022demes}
encompassing hundreds of populations, explosive growth rates and
myriad interconnections of migration, were somehow estimated and
provided as input, ARGs sampled from
the coalescent cannot capture the complexities of realistic family structures, which often involve overlapping generations and extreme variation in lifetime reproductive success. Flexible ARG encoding should accommodate such cases.

>>>>

Fully decoupling the general concept of an ARG from the coalescent
with recombination stochastic process is an important step.
While this has proven to be a useful and
robust
model~\citep{wakeley2012gene,bhaskar2014distortion,nelson2020accounting},
many modern datasets have properties that grossly
violate its assumptions.
For instance, several human datasets now consist of hundreds of thousands of
genomes~\citep{turnbull2018hundred, bycroft2018genome,karczewski2020mutational,tanjo2021practical},
and so sample size is substantially \emph{larger} than the
usually assumed $N_e$ values.
Agricultural datasets are an even more extreme departure from coalescent
assumptions, with hundreds of thousands of samples embedded in
multi-generational pedigrees~\citep{hayes20191000,Ros-Freixedes2020}
and effective population sizes of 100 and even
less~\citep{MacLeod2013,Makanjuola2020,Hall2016,Porcnic2016}.
Even if sufficiently complex demographic models~\citep{gower2022demes}
encompassing hundreds of populations, explosive growth rates and
myriad interconnections of migration, were somehow estimated and
provided as input, ARGs sampled from
the coalescent cannot capture the complexities of realistic family structures, which often involve overlapping generations and extreme variation in lifetime reproductive success. Flexible ARG encoding should accommodate such cases.
